### 2019-05-20 ch1_singlecellbiology.tex
initial commit

\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{Single Cell Biology} 
\chaptermark{Single Cell Biology.} % short title
\label{chap:singlecellbiology}
\lettergroup{\thechapter}

\begin{quote}
	\textbf{Abstract:} \blindtext
\end{quote}

\vfill

Adapted from:\\
\textbf{Cannoodt, R.}$^*$, Saelens, W.$^*$, and Saeys, Y. Computational methods for trajectory inference from single-cell transcriptomics. \textit{European Journal of Immunology} 46, 11 (2016), 2496--2506. \doi{10.1002/eji.201646347}.\\
{\footnotesize $^*$ Equal contribution}
\newpage

\blindtext