### 2019-05-20 ch3_scorpius.tex

\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{SCORPIUS} 
\chaptermark{SCORPIUS.} % short title
\label{chap:scorpius}

\begin{quote}
	\textbf{Abstract:} \blindtext
\end{quote}

\vfill

Adapted from:\\
\textbf{Cannoodt, R.}, \ldots, De Preter, K., and Saeys, Y. SCORPIUS improves trajectory inference and identifies novel modules in dendritic cell development. \textit{Journal} vol, issue (2019), page--page. \doi{10.1101/079509v2}.
\newpage

\blindtext