### 2019-05-22 ch2_dynbenchmark.tex
update document

\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{A comparison of single-cell trajectory inference methods} 
\chaptermark{A comparison of single-cell trajectory inference methods.} % short title
\label{chap:dynbenchmark}

\begin{quote}
	\textbf{Abstract:} \blindtext
\end{quote}

\vfill

Adapted from:\\
Saelens, W.$^*$, \textbf{Cannoodt, R.}$^*$, Todorov, H., and Saeys, Y. A comparison of single-cell trajectory inference methods. \textit{Nature Biotechnology} 37, 5 (2019), 547--554. \doi{10.1038/s41587-019-0071-9}.\\
{\footnotesize $^*$ Equal contribution}
\newpage

\blindtext