### 2019-05-22 ch6_incgraph.tex

\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{Optimising regulatory networks} 
\chaptermark{Optimising regulatory networks.} % short title
\label{chap:incgraph}

\begin{quote}
	\textbf{Abstract:} \blindtext
\end{quote}

\vfill

Adapted from:\\
\textbf{Cannoodt, R.}, Ruyssinck, J., Ramon, J., De Preter, K., and Saeys, Y. IncGraph: Incremental graphlet counting for topology optimisation. \textit{PLOS ONE} 13, 4 (2018), e0195997. \doi{10.1371/journal.pone.0195997}.
\newpage

\blindtext \cite{Eren2013}