### 2019-08-11 ch0_frontmatter.tex
fix logos

\hypersetup{%
	pdftitle = {Modelling single cell dynamics with trajectories and gene regulatory networks},
	pdfsubject = {A subject},
	pdfkeywords = {import keywords},
	pdfauthor = {Robrecht Cannoodt},
	pdfcreator = {\LaTeX},
}

%%%%%%%%%%%%%%%%%%%%%%%%
%%% DEFINE TITLEPAGE %%%
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{titlepage}
	\centering
	\ 
	\vfill
	{\Huge \textsf{Modelling single cell dynamics\\with trajectories and gene regulatory networks}}

	\vspace{2.25cm}

	{\LARGE \textsf{Robrecht Cannoodt}}

	\vspace{2.25cm}

	{\textsf{Thesis submitted in fulfilment of the requirements for the degree of \\Doctor in Computer Science, 2019}}

	\vfill

	{\textsf{Supervisors: \\
			Prof. Dr. Yvan Saeys\\Prof. Dr. Katleen De Preter}}

	\vfill

	\includegraphics[height=.7cm]{fig/logos/icon_UGent_WE_EN_RGB_color}\hfill
	\includegraphics[height=1cm]{fig/logos/vib_rf_inflammation_research_rgb_pos} \hfill
	\includegraphics[height=.8cm]{fig/logos/dambi_logo} \hfill
	\includegraphics[height=.7cm]{fig/logos/cmgg_myshort} \hfill
	\includegraphics[height=.7cm]{fig/logos/FWO_Logo_KleurCMYK} 
	\vspace{1cm}

	{\textsf{Vakgroep Toegepaste Wiskunde, Informatica, en Statistiek\\
			Faculteit Wetenschappen, Universiteit Gent \\
			Krijgslaan 281 - S2, 9000 Gent}}

\end{titlepage}

%%%%%%%%%%%%%%%%%%%%%%%%
%%% ACKNOWLEDGEMENTS %%%
%%%%%%%%%%%%%%%%%%%%%%%%

\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{Acknowledgements}
\blindtext

\newpage{\thispagestyle{empty}\cleardoublepage}

%%%%%%%%%%%%%
%%% LISTS %%%
%%%%%%%%%%%%%
\newpage{\thispagestyle{empty}\cleardoublepage}
\tableofcontents

%\newpage{\thispagestyle{empty}\cleardoublepage}
%\listoffigures

%\newpage{\thispagestyle{empty}\cleardoublepage}
%\listoftables

\newpage{\thispagestyle{empty}\cleardoublepage}

%%%%%%%%%%%%%%%%%%%%
%%% NOMENCLATURE %%%
%%%%%%%%%%%%%%%%%%%%
\makenomenclature
\printnomenclature

\newcommand{\grn}{GRN}
\newcommand{\grns}{GRNs}
\nomenclature{GRN}{Gene Regulatory Network}

\newcommand{\im}{IM}
\newcommand{\ims}{IMs}
\nomenclature{IM}{Importance Measure}

\newcommand{\tf}{TF}
\newcommand{\tfs}{TFs}
\nomenclature{TF}{Transcription Factor}

\newcommand{\dna}{DNA}
\nomenclature{DNA}{Deoxyribonucleic Acid}

\newcommand{\rna}{RNA}
\nomenclature{RNA}{Ribonucleic Acid}

\newcommand{\mrna}{mRNA}
\nomenclature{mRNA}{Messenger RNA}

\newcommand{\nin}{NI}
\nomenclature{NI}{Network Inference}

\newcommand{\ml}{ML}
\nomenclature{ML}{Machine Learning}

\newcommand{\rf}{RF}
\nomenclature{RF}{Random Forests}

\newcommand{\cart}{CART}
\nomenclature{CART}{Classification And Regression Trees}