### 2019-08-23 ch2_dyngen.tex
update

\newpage{\thispagestyle{empty}\cleardoublepage}
\chapter{dyngen: Simulating single-cell data using gene regulatory networks} 
\chaptermark{dyngen: Simulating single-cell data using gene regulatory networks} % short title
\label{chap:dyngen}

\begin{quote}
	\textbf{Abstract:} 
\end{quote}

\vfill

Adapted from:\\

\newpage

\begin{figure}
	\centering
	\includegraphics[width=\LARGEfigure]{fig/dyngen/showcase_2.png} 
	\caption{Showcase of dyngen functionality. \textbf{TODO: change to pdf.}} % TODO: update label
	\label{fig:dyngen}
\end{figure}