### 2019-08-26 ch8_discussion/content.tex
refactorise

\blindtext

\section{Overview and conclusions of the presented work}

% notice that normalisation was absent from this dissertation.
%Computer scientists should proceed with caution in developing new software, however, as the results produced should not only be convincing and easy to interpret, but the software should be robust and generate sufficiently accurate models of the underlying system.
%
%Bit too excited -- false positives, poor accuracy, scalability issues, poor software quality.

\section{Future research directions}