### 2019-09-27 ch5b_incgraph/fig/pseudocodeS1.tex
update

\documentclass[a4paper,11pt]{article}

\usepackage[scaled]{helvet}
\renewcommand\familydefault{\sfdefault} 
\usepackage[T1]{fontenc}

\newcommand{\pluseq}{\mathrel{+}=}
\newcommand{\asteq}{\mathrel{*}=}

% Drawing with TiKz
\usepackage{tikz}
\usetikzlibrary[positioning]
\usetikzlibrary{patterns}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{shapes.arrows}
\usetikzlibrary{arrows}
\usetikzlibrary{calc,positioning,shapes,chains,mindmap,trees,decorations}
\usetikzlibrary{matrix,fit}

\usepackage{algorithm}
\usepackage{fixltx2e}
\usepackage{algpseudocode}
\MakeRobust{\Call}

\usepackage[small, bf]{caption}
\usepackage{subcaption}

\usepackage{float}

\usepackage[cm]{fullpage}

\usepackage{pdflscape}

\newcommand{\subsubsec}[1]{{\setlength{\parindent}{0cm}\textbf{#1} \\}}
\newcommand{\specialcell}[2][c]{\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\newcommand\mycaption[4]{%
  \textbf{#1:\ \ \ #3} #4}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\pagestyle{empty}
\begin{document}

\begin{algorithmic}
	\Function{CalculateDelta}{Network $G'$, Node $n_0$, Node $n_1$}
		\State \Comment{Assuming an edge $(n_0, n_1)$ has just been added or removed from $G'$}
		\State $\Delta^- \gets \Call{Matrix}{\Call{NumNodes}{G}, 73}$ \Comment{For storing the orbit counts of the removed graphlets}
		\State $\Delta^+ \gets \Call{Matrix}{\Call{NumNodes}{G}, 73}$ \Comment{For storing the orbit counts of the added graphlets}
		\State $B_0 = \{n_0, n_1\}$ \Comment{A blacklist of nodes not to visit anymore}
		\State $e = (n_0, n_1)$ \Comment{Different name for the edge}
		\If{$e \in N'$}
			\State $(m^-, m^+) = (0, 1)$ \Comment{In other words, $m^- = 1$ iff $e \in N$ and $m^+ = 1$ iff $e \in N'$}
		\Else
			\State $(m^-, m^+) = (1, 0)$
		\EndIf
		\State $x_0 = 0$
		\State $(\Delta^+, \Delta^-) = \Call{CountOrbits}{\Delta^+, \Delta^-, (n_0, n_1), x_0, m^-, m^+}$ \Comment{Update delta matrices for current nodes}

		%\For{$n_2 \in \Call{Neighbours}{S_0}$}
		\For{$n_2 \in \bigcup_{i \in \{0, 1\}}\Call{Neighbours}{n_i}$}
			\If{$n_2 \not\in B_0$}
				\State $B_0 = \{n_2\} \bigcup B_0$ \Comment{Add $n_2$ to blacklist $B_0$}
				\State $x_1 = x_0 + \Call{W}{G', (n_0, n_2), 1} + \Call{W}{G', (n_1, n_2), 2}$ \Comment{Calculate edge weights for current nodes}
				\State $(\Delta^+, \Delta^-) = \Call{CountOrbits}{\Delta^+, \Delta^-, (n_0, n_1, n_2), x_1, m^-, m^+}$
				\State $B_1 = B_0$ \Comment{Make a copy for the next iteration depth}
				\For{$n_3 \in \bigcup_{i \in \{0, 1, 2\}}\Call{Neighbours}{n_i}$}
					\If{$n_3 \not\in B_1$}
						\State $B_1 = \{n_3\} \bigcup B_1$
						\State $x_2 = x_1 + \Call{W}{G', (n_0, n_3), 3} + \Call{W}{G', (n_1, n_3), 4} + \Call{W}{G', (n_2, n_3), 5}$
						\State $(\Delta^+, \Delta^-) = \Call{CountOrbits}{\Delta^+, \Delta^-, (n_0, n_1, n_2, n_3), x_2, m^-, m^+}$
						\State $B_2 = B_1$

						\For{$n_4 \in \bigcup_{i \in \{0, 1, 2, 3\}}\Call{Neighbours}{n_i}$}
							\If{$n_4 \not\in B_2$}
								\State $B_2 = \{n_4\} \bigcup B_2$
								\State $x_3 = x_2 + \Sigma_{i \in \{0, 1, 2, 3\}}\Call{W}{G', (n_i, n_4), i+6}$
								\State $(\Delta^+, \Delta^-) = \Call{CountOrbits}{\Delta^+, \Delta^-, (n_0, n_1, n_2, n_3, n_4), x_3, m^-, m^+}$
							\EndIf
						\EndFor
					\EndIf
				\EndFor
			\EndIf
		\EndFor
	\EndFunction
	\Function{CountOrbits}{$\Delta^+$, $\Delta^-$, Nodes $S$, Edgeweights $x$, Modifier $m^-$, Modifier $m^+$}
  \State $L^- = L[x + m^-]$ \Comment{Look up the orbits of the subgraph induced by $S$ in $N$}
			\State $\Delta^-[S, L^-] \pluseq 1$ \Comment{Increment orbit counts of nodes $S$ at positions $L^-$ in $\Delta^-$}
  \State $L^+ = L[x + m^+]$ \Comment{Look up the orbits of the subgraph induced by $S$ in $N'$}
			\State $\Delta^+[S, L^+] \pluseq 1$ \Comment{Increment orbit counts of nodes $S$ at positions $L^+$ in $\Delta^+$}
			\State \Return $(\Delta^-, \Delta^+)$
	\EndFunction
	\Function{W}{Network $G$, Edge $e$, Exponent $i$}
		\State \Return $e \in \Call{Edges}{G} \ ?\  2^i : 0$ \Comment{Return $2^i$ if $G$ contains edge $e$}
	\EndFunction
\end{algorithmic}

\end{document}