### 2019-09-29 ch7_incgraph/fig/pseudocodeS4.tex
restructure

\documentclass[a4paper,11pt]{article}

\usepackage[scaled]{helvet}
\renewcommand\familydefault{\sfdefault} 
\usepackage[T1]{fontenc}

\newcommand{\pluseq}{\mathrel{+}=}
\newcommand{\asteq}{\mathrel{*}=}

% Drawing with TiKz
\usepackage{tikz}
\usetikzlibrary[positioning]
\usetikzlibrary{patterns}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{shapes.arrows}
\usetikzlibrary{arrows}
\usetikzlibrary{calc,positioning,shapes,chains,mindmap,trees,decorations}
\usetikzlibrary{matrix,fit}

\usepackage{algorithm}
\usepackage{fixltx2e}
\usepackage{algpseudocode}
\MakeRobust{\Call}

\usepackage[small, bf]{caption}
\usepackage{subcaption}

\usepackage{float}

\usepackage[cm]{fullpage}

\usepackage{pdflscape}

\newcommand{\subsubsec}[1]{{\setlength{\parindent}{0cm}\textbf{#1} \\}}
\newcommand{\specialcell}[2][c]{\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\newcommand\mycaption[4]{%
  \textbf{#1:\ \ \ #3} #4}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\pagestyle{empty}
\begin{document}

	\begin{algorithmic}
		\Function{Geometric}{number of nodes $n$, number of edges $e$, \par\hspace*{2.77cm} number of operations $o$, number of dimensions $d=3$}
			\State $P \gets$ Sample $n$ points from a multivariate continuous uniform distribution $U_{d}((0,1)^{d})$
			\State $D \gets$ Calculate distance matrix from $P$
			\State $N_0 \gets \Call{Head}{\Call{ArgSort}{\Call{LowerTriangle}{D}}, e}$ \Comment{Initial network}
			\State $N_p \gets N_0$
			\State $O \gets (\,)$ \Comment{Variable to store the generated operations in}
			\While{$|O| < o$} \Comment{Modify network until $O$ is sufficiently large}
				\State $i \gets $ Sample 1 index from $\{0\,..\,n-1\}$
				\State $P[i] \gets$ Sample 1 point from a $U_{d}((0,1)^{d})$ \Comment{Give node $i$ a new location}
				\State $D[i,] \gets D[,i] \gets $ Calculate new distances between node $i$ and all other nodes
				\State $N_c \gets \Call{Head}{\Call{ArgSort}{\Call{LowerTriangle}{D}}, e}$ \Comment{New network}
				\State $O_a \gets \{(\textrm{ADD}, e) \,|\, e \not\in E(N_p) \wedge e \in E(N_c) \}$ \Comment{Gather added edges}
				\State $O_r \gets \{(\textrm{REM}, e) \,|\, e \in E(N_p) \wedge e \not\in E(N_c) \}$ \Comment{Gather removed edges}
				\State $O \gets O + \Call{Shuffle}{O_a + O_r}$ \Comment{Append new operations to $O$}
				\State $N_p \gets N_c$
			\EndWhile
			\State $O \gets \Call{Head}{O, o}$ 
			\State \Return $(N_0, O)$ \Comment{Return the initial geometric network and $o$ operations}
		\EndFunction
	\end{algorithmic}

\end{document}