### 2019-11-22 defence_internal/pres_v1.tex
~

\documentclass[usepdftitle=false, compress]{beamer}

\usepackage{calc}
\newlength{\leftstart}
\newlength{\topstart}
\setlength{\leftstart}{\hoffset + \oddsidemargin + 1in}
\setlength{\topstart}{\voffset + \topmargin + 1in}

\newlength{\figfour}

\usepackage{etex}
\usepackage{etoolbox}
\usepackage[absolute,overlay]{textpos}
\usepackage{fancybox}

\usepackage{xcolor}
\definecolor{myblue}{rgb}{0,0.4,0.7}
\definecolor{myred}{rgb}{0.7,.1,.1}
\definecolor{mygray}{rgb}{0.7,0.7,0.7}
\definecolor{mygreen}{rgb}{0,0.8,0.1}

\useinnertheme{circles}

\DeclareMathOperator*{\argmin}{arg\,min}

\mode<presentation>{
  \setbeamertemplate{navigation symbols}{}
	\usepackage{lmodern}
	\setbeamercolor{title}{fg=myblue}
	\setbeamercolor{frametitle}{fg=black}
}

\newcommand{\myemph}[1]{{\color{myblue}#1}}

% Drawing with TiKz
\usepackage{tikz}
\usetikzlibrary[positioning]
\usetikzlibrary{patterns}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{shapes.arrows}
\usetikzlibrary{arrows}
\usetikzlibrary{calc,positioning,shapes,chains,mindmap,trees,decorations}

\usepackage{rotating}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amssymb,amsmath,amsthm}
\usepackage{graphicx,graphics}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{tabularx}
\newcolumntype{Y}{>{\centering\arraybackslash}X}
\newcolumntype{Z}{>{\raggedright\arraybackslash}X}
\newcolumntype{W}{>{\raggedleft\arraybackslash}X}

\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\usepackage{pgfgantt}

\usepackage[pdf]{pstricks}

\usepackage[absolute,overlay]{textpos}

\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\color{mygreen}\ding{51}}%
\newcommand{\xmark}{\color{myred}\ding{55}}%
\newcommand{\mehhh}{\color{mygray}$\sim$}%

\newcommand{\prog}{ $\vartriangleright$ }

\usepackage{enumitem}
\setitemize{label=\usebeamerfont*{itemize item}%
  \usebeamercolor[fg]{itemize item}
  \usebeamertemplate{itemize item}}

\newenvironment{compitem}{\begin{itemize}[noitemsep,topsep=0pt,parsep=0pt,partopsep=0pt]}{\end{itemize}}

\newcommand{\wholeslide}[1]{{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{#1}}
\begin{frame}[plain]
\end{frame}
}}
\usepackage{soul}

\makeatletter
\def\imod#1{\allowbreak\mkern7mu{\operator@font mod}\,\,#1}
\def\m#1{\allowbreak\mkern10mu({\operator@font mod}\,\,#1)}
\makeatother

\newcommand{\summary}[5]{%
	\begin{frame}
		\begin{itemize}
			\item {\color{#1}Single-cell biology}
			\item {\color{#2}Trajectory inference}
			\item {\color{#3}Network inference}
			\item {\color{#4}Benchmarking}
			\item {\color{#5}Practical implications}
		\end{itemize}
	\end{frame}
}

\title{Modelling single-cell dynamics\\with trajectories and gene regulatory networks}

\author{Robrecht Cannoodt}

\date{}

\begin{document}

\begin{frame}[plain]
  \titlepage
  \hfill
  \includegraphics[height=.42cm]{../fig/logos/icon_UGent_WE_EN_RGB_color}\hfill
  \includegraphics[height=.60cm]{../fig/logos/vib_rf_inflammation_research_rgb_pos} \hfill
  \includegraphics[height=.48cm]{../fig/logos/dambi_logo} \hfill
  \includegraphics[height=.42cm]{../fig/logos/cmgg_myshort} \hfill
  \includegraphics[height=.42cm]{../fig/logos/FWO_Logo_KleurCMYK} 
  \hfill
\end{frame}

\begin{frame}
	\frametitle{Single-cell biology \uncover<2->{is to study an organism}\\\uncover<3->{by studying the behaviour of its cells}}
	\begin{center}
		\only<2>{\includegraphics[width=\linewidth]{figures/1a_singlecellbiology}}%
		\only<3>{\includegraphics[width=\linewidth]{figures/1b_singlecellbiology}}%
		\only<4>{\includegraphics[width=\linewidth]{figures/1c_singlecellbiology}}%
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Technological advances allow high-throughput\\single-cell biology}
	\begin{center}
		\includegraphics[width=\linewidth]{figures/2_part_scaling.pdf}\\
		\flushright 		{\vspace*{-1em}\footnotesize Regev et al. 2018}
	\end{center}
  \vspace*{-2.2em}
  \uncover<2->{
	  \rotatebox[origin=c]{180}{$\Lsh$} Trajectory inference \\
	  \rotatebox[origin=c]{180}{$\Lsh$} Network inference \\
	}
\end{frame}

%%%%%%%%%%%%%% SUMMARY
\summary{lightgray}{black}{lightgray}{lightgray}{lightgray}

\begin{frame}
	\frametitle{Cells are highly dynamic entities}
	\begin{center}
		\includegraphics[width=.6\linewidth]{figures/3_dynamiccells.pdf}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{A typical trajectory inference analysis}
	\begin{center}
		\only<1>{\includegraphics[width=\linewidth]{figures/4a_trajectoryinference.pdf}}%
		\only<2>{\includegraphics[width=\linewidth]{figures/4b_trajectoryinference.pdf}}%
		\only<3>{\includegraphics[width=\linewidth]{figures/4c_trajectoryinference.pdf}}%
		\only<4>{\includegraphics[width=\linewidth]{figures/4d_trajectoryinference.pdf}}%
		\only<5>{\includegraphics[width=\linewidth]{figures/4e_trajectoryinference.pdf}}%
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Benchmark of 45 trajectory inference methods}
	\begin{center}
		\only<1>{\includegraphics[width=.9\linewidth]{figures/5a_benchmark.pdf}}%
		\only<2>{\includegraphics[width=.9\linewidth]{figures/5b_benchmark.pdf}}%
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{Interactive guidelines}
	\begin{center}
		\includegraphics[width=\linewidth]{figures/6_guidelines.pdf}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{SCORPIUS: Linear trajectory inference}
	\begin{center}
		\includegraphics[width=.8\linewidth]{figures/end_part_scorpius.pdf}
	\end{center}
\end{frame}

\begin{frame}
	\frametitle{dyno: A toolkit for inferring and interpreting trajectories}
	\begin{center}
		\includegraphics[width=\linewidth]{figures/end_part_dyno.pdf}
	\end{center}
\end{frame}
\begin{frame}
	\frametitle{Perspective: self-assessment in trajectory inference}
	\begin{center}
		\includegraphics[width=.8\linewidth]{figures/end_part_selfassessment.pdf}
	\end{center}
\end{frame}

%%%%%%%%%%%%%% SUMMARY
\summary{lightgray}{lightgray}{black}{lightgray}{lightgray}

%%%%%%%%%%%%%% SUMMARY
\summary{lightgray}{lightgray}{lightgray}{black}{lightgray}

\begin{frame}
	\frametitle{dyngen}
	\begin{center}
		\includegraphics[width=.6\linewidth]{figures/end_part_dyngen.pdf}
	\end{center}
\end{frame}

\begin{frame}
\frametitle{Guidelines for benchmarking computational tools}
\begin{center}
	\includegraphics[width=\linewidth]{figures/end_part_guidelines.pdf}
\end{center}
\end{frame}

%%%%%%%%%%%%%% SUMMARY
\summary{lightgray}{lightgray}{lightgray}{lightgray}{black}

\end{document}