### 2019-12-10 ack.tex
update

Hello there! Did you stumble across this phrase by accident, 
or are you looking for what I think you're looking for?
For every page you turn over in this book,
I feel that there is at least one person who deserves to be thanked 
a thousand times over for their continued love, sacrifices and support.
So here we go, until the end of the book: Thank you! Thank you!
Thank you! Thank you! Thank you! Thank you! Thank you! Thank you! 
Thank you! Thank you! Thank you! Thank you! Thank you! Thank you!
Thank you! Thank you! Thank you! Thank you! Thank you! Thank you!
Thank you! Thank you! Thank you! Thank you! Thank you! Thank you!
Thank you! Thank you! Okay, enough with the silly business.
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Yvan, you are an incredible supervisor and group leader. 
You always tell the baby PhD students that the first step in a PhD
is not learning to write well, but learning to like to write.
Those words are amazing PhD training wheels.
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Katleen, thank you for your dedication and perseverance
in guiding me and making sure I got off to a good start in my PhD.
I see exciting times ahead for the KDP group, which is overflowing
with enthusiasm and inspiration. I wonder what the future will bring!
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Frank, thank you for giving me the opportunity to start on this journey.
You are an inexhaustible source of new ideas
which often explore the very thin line
between ingenuity and completely insane.
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Jo, sorry for spilling coffee on your shoes on my first day in Blok B.
My strategy seemed to have worked, for suddenly I had a fourth promotor!
I thoroughly enjoyed my time as an unofficial member of your group.
You are a true people's person; you bring out the best in everybody.
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Joeri, we've come full circle. You started me on this adventure
with your incredible friendliness and that darned oversized pumpkin;
and now, with you as my jury member, you helped me reach my goal.
Let's stay in touch :)
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Wouter, I cannot imagine a better partner to do a PhD with.
If we were genes, I'm sure we would reciprocal activators,
constantly stimulating each other to do just a little bit more!
Yet, our knowledge was often so perfectly complementary;
if we were a PCA plot, you'd be PC1 and I'd be PC2.
Or maybe the other way around? Let's take turns being PC1 and PC2.
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Sofie, you saved my rear-end from embarrassing moments more frequently than I dare to admit.
The perfect example was when I walked in the MMM on 2014-01-06,
I arrived a little bit too late, and you send me a message:
"You do know that you have to present today, right?"
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Helena, thank you for being such a lovely colleague and an even better friend.
I enjoy our scientific talks and our non-scientific walks. 
Together with you, I played the longest ongoing game of hide-and-seek,
on page 22-9 you probably already found the truc that I am referring to.
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Paulina, like a soft summer wind you passed through our lives.
Thanks for putting so many healthy vege-tables and strange Polish delicacies in our fridge,
for cycling together to work, and for making weird noises at Remi.
I'm looking forward to the next time we meet, so you can be amazed at how much Remi has grown.
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Soup: Dani, Helena, Caroline, Robin B, Maxim, Annelies, Quentin, Katrien, Arne G, Ruth, Joris
Dambi
Dambi
Dambi
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Miriam
Miriam
Miriam
Miriam
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Laure, Michael, Jasper en Lize (spelling??)
Mara, Seppe \& Alice
Marta, Jan, Tristan, Rafael
Daan, Mieke \& Lou
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Cedric
Cedric
Cedric
Cedric
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Familie
Familie
Familie
Familie
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Victor, humor
Victor
Victor
Victor
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Hendrik, getting into PC
Hendrik 
Hendrik
Hendrik
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Papa Chris, babbel voor het slapen gaan, de vele schaakspelletjes
Chris
Chris
Chris
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Mama en papa Jean-Paul: opvoeding, zorgeloosheid
Mama
Mama
Mama
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Amor, as I write this, you are sitting on my lap and drooling on the table.
I guesstimate about 50\% of this dissertation was written like this.
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Hello Petit Remi! 
Remi
Remi
Remi
\rule{10cm}{0.4pt}
\rule{10cm}{0.4pt}
Dear Caro,
I cannot imagine a better partner and mother than you.
You are a strange combination of being extremely thoughtful and caring,
yet you roll over laughing when I stub my toe against the side of the table.
Let's get back on track collecting doorsteps,
and spending time doing the things we love, just us three.
I look forward to the rest of our lives together,
my strong independent woman.